Analysis and calculations in manuscript

### 2017-04-06 EU-Circle/eu-circle_case-study-2.Rmd
Cleansed version

@@

```{r,warning=FALSE}
if (!file.exists('eu-circle_case-study-2_slp.rda')) {
  x <- retrieve('pp_0.25deg_reg_v13.1.nc',path='~/Data/data.EOBS',
                lon=c(10,30),lat=c(53,66))
  save(file='eu-circle_case-study-2_slp.rda',x)
} else load('eu-circle_case-study-2_slp.rda')
## Only keep the ECAD data
map(x,FUN='q95',new=TRUE)
diagnose(x)
```

@@

```{r,warning=FALSE}
if (!file.exists('eu-circle_case-study-2_slp.rda')) {
  x <- retrieve('pp_0.25deg_reg_v13.1.nc',path='~/Data/data.EOBS',
                lon=c(10,30),lat=c(53,66))
  save(file='eu-circle_case-study-2_slp.rda',x)
} else load('eu-circle_case-study-2_slp.rda')
## Only keep the ECAD data
map(x,FUN='q95',new=TRUE)
diagnose(x)
```

**Ship accidents in the Baltic Sea**
baltic.events <- c(19931010,20000328,20010328,20040313,20060330,20081017,20081202,20091002,20100128)

>>>>

Synoptic storms are associated with winds, storm surges and waves.

```{r}
load('cyclones.baltic.rda')
map(cyclones.baltic)
```

Plot the number of trajectories over time

```{r}
plot(cyclones.baltic)
```

@@

```{r,warning=FALSE}
if (!file.exists('eu-circle_case-study-2_slp.rda')) {
  x <- retrieve('pp_0.25deg_reg.nc',path='~/storeB/data/data.ECAD',
                lon=c(10,30),lat=c(53,66))
  save(file='eu-circle_case-study-2_slp.rda',x)
} else load('eu-circle_case-study-2_slp.rda')
## Only keep the ECAD data
map(x,FUN='q95',new=TRUE)
diagnose(x)
```

@@

**Ship accidents in the Baltic Sea**

```{r}
baltic.events <- c(19931010,20000328,20010328,20040313,20060330,20081017,20081202,20091002,20100128)
uv <- geostrophicwind(x)
```

### 2017-04-06 EU-Circle/eu-circle_case-study-3.Rmd
Cleansed version

@@

```{r predictor, warning=FALSE, message=FALSE}
## The mean sea-level pressure
slp <- retrieve('slp.mon.mean.nc',lon=c(-80,0),lat=c(20,70))
## The surface temperature
t2m <- retrieve('air.mon.mean.nc',lon=c(-80,0),lat=c(10,70))
## Saturation water pressure estimated from maritime temperatures
es <- C.C.eq(mask(t2m,land=TRUE))
eof.slp<- EOF(annual(slp))
eof.t2m <- EOF(annual(t2m))
eof.es <- EOF(annual(es,FUN='mean'))
```

@@

```{r newlyn, results="hide", warning=FALSE, message=FALSE}
newlyn <- station.newlyn()
newlyn <- aggregate(newlyn,year,FUN='max',na.rm=TRUE)
```

@@

```{r combine-sea-level, warning=FALSE, message=FALSE}
index(newlyn) <- year(newlyn)
index(sonel) <- year(sonel)
index(gloss) <- year(gloss)
Z <- combine.stations(newlyn,sonel,gloss)
nv <- apply(coredata(Z),2,FUN='nv')
Z <- subset(Z,is=(nv > 50))
Z <- subset(Z,it=c(1950,2015))
plot(Z); grid()
```

@@

```{r sl-trend-map, warning=FALSE, message=FALSE}
zc <- coredata(Z)
zc[is.infinite(zc)] <- NA
zc -> coredata(Z)
map(Z,FUN='trend',cex=-2)
```

@@

```{r zonal-wind-check-get-slp, warning=FALSE, message=FALSE}
## Trend in upper wind direction
SLP <- retrieve('slp.mon.mean.nc',lon=c(-50,30),lat=c(40,60))
eof.slp<- EOF(annual(SLP))
```

@@

```{r reanalysis-wind-speed, warning=FALSE, message=FALSE}
## Estimate grostrophich wind
UV <- geostrophicwind(SLP,progressbar=FALSE)
```

```{r wind-rose}
## wind rose
windrose(uv)
```

>>>>

```{r, eval=FALSE}
## Get files on the outgoing long-wave radiation (OLR).
cmip5.urls <- function(experiment='rcp45',varid='rlut',
                       url="http://climexp.knmi.nl/CMIP5/monthly/", 
                       path=NULL,off=FALSE,force=FALSE,verbose=FALSE) {
  urlfiles <- "NA"
  if(verbose) print("cmip5.urls")
  if(is.null(path)) path <- getwd()
  for (iexp in experiment) {
    if(verbose) print(iexp)
    for (ivar in varid) {
      if(verbose) print(ivar)
      ## Loop on the number of experiments
      for (irun in 0:110) { ## 
        if(verbose) print(paste(irun))
        ## Update experiment number
        if (irun < 10) run.id = paste("00",as.character(irun),sep="")
        else if (irun < 100) run.id = paste("0",as.character(irun),sep="")
        else run.id <- as.character(irun)

        ## Create output directory for the climate experiment
        path.exp <- file.path(path,experiment[grep(iexp,experiment)],
                              fsep = .Platform$file.sep)
        if (!file.exists(path.exp)) dir.create(path.exp)
        if (verbose) print(path.exp[grep(iexp,experiment)])
        ## Define the output file
        destfile <- paste(path.exp,varid[grep(ivar,varid)],sep="/") 
        destfile <- paste(destfile,"_Amon_ens_",sep="")
        destfile <- paste(destfile,iexp,sep="")
        destfile <- paste(destfile,run.id,sep="_")
        destfile <- paste(destfile,".nc",sep="")

        if (!file.exists(destfile) | force) {
          ## Update output filename with attributes:
          urlfile  <- paste(url,ivar,sep="")             # add var directory
          urlfile  <- paste(urlfile,ivar,sep="/")        # add v.name
          urlfile  <- paste(urlfile,"_Amon_ens_",sep="") # add text
          urlfile  <- paste(urlfile,iexp,sep="")         # add exp.name
          urlfile  <- paste(urlfile,run.id,sep="_")      # add exp ID number
          urlfile  <- paste(urlfile,".nc",sep="")        # add file ext
        }
        urlfiles <- c(urlfiles,urlfile)
        if (verbose) print(urlfile)
      }
    } # End for   
  }
  return(urlfiles[-1])
}

## download the CMIP5-files from the KNMI ClimateExplorer
urls <- cmip5.urls()
for (i in 1:length(urls)) if (!file.exists(substr(urls[i],43,70))) download.file(urls[i],substr(urls[i],43,70))

**Wind**

Synoptic storms are associated with winds, storm surges and waves.

```{r}
load('cyclones.uk.rda')
map(cyclones.uk,new=FALSE)
```

Plot the number of trajectories over time

```{r}
plot(cyclones.uk,new=FALSE)
```

@@

```{r predictor, warning=FALSE, message=FALSE}
## The mean sea-level pressure
slp <- retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-80,0),lat=c(20,70))
## The surface temperature
t2m <- retrieve('~/Downloads/air.mon.mean.nc',lon=c(-80,0),lat=c(10,70))
## Saturation water pressure estimated from maritime temperatures
es <- C.C.eq(mask(t2m,land=TRUE))
eof.slp<- EOF(annual(slp))
eof.t2m <- EOF(annual(t2m))
eof.es <- EOF(annual(es,FUN='mean'))
```

@@

```{r newlyn, results="hide", warning=FALSE, message=FALSE}
if (!file.exists('newlyn.rda')) {
  newlyn <- station.newlyn()
  newlyn <- aggregate(newlyn,year,FUN='max',na.rm=TRUE)
  save(newlyn,file='newlyn.rda')
} else load('newlyn.rda')
```

@@

```{r combine-sea-level, warning=FALSE, message=FALSE}
index(newlyn) <- year(newlyn)
index(sonel) <- year(sonel)
index(gloss) <- year(gloss)
Z <- combine.stations(newlyn,sonel,gloss)
nv <- apply(coredata(Z),2,FUN='nv')
Z <- subset(Z,is=(nv > 50))
Z <- subset(Z,it=c(1966,2015))
coredata(Z)[!is.finite(Z)] <- NA
coredata(Z) <- t(t(coredata(Z)) - colMeans(coredata(Z),na.rm=TRUE))
plot(Z); grid()
```

@@

```{r sl-trend-map, warning=FALSE, message=FALSE, fig.height=8}
zc <- coredata(Z)
zc[is.infinite(zc)] <- NA
zc -> coredata(Z)
map(Z,pch=19,FUN='trend',cex=-2,add.text=TRUE)
```

Get the annual mean global mean sea-level and analyse the correspondense between that and the local sea level - we can expect a link from heuristic physics. Also include an account of the mean sea-level pressure.

```{r, warning=FALSE, message=FALSE, fig.height=8}
gsl <- annual(GSL())
plot(gsl)
slp <- annual(retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-20,10),lat=c(45,60)))
eof.slp <- EOF(slp,n=5)
## Manipulate the data: replace the last SLP EOF with the global sea-level and use DS to explore connections
i1 <- is.element(year(eof.slp),year(gsl))
i2 <- is.element(year(gsl),year(eof.slp))
coredata(eof.slp)[i1,5] <- coredata(gsl)[i2]
attr(eof.slp,'pattern')[,,5] <- 1
attr(eof.slp,'eigenvalue') <- 1
plot(eof.slp,ip=5,new=FALSE)
```

Prepare the local tidal data: PCA

```{r, warning=FALSE, message=FALSE, fig.height=8}
nv <- apply(coredata(Z),2,'nv')
Z <- subset(Z,is=nv > 45)
Z <- pcafill(Z)
pca.Z <- PCA(Z)
plot(pca.Z, new=FALSE)
```

@@

```{r zonal-wind-check-get-slp, warning=FALSE, message=FALSE}
## Trend in upper wind direction
SLP <- retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-50,30),lat=c(40,60))
eof.slp<- EOF(annual(SLP))
```

@@

```{r reanalysis-wind-speed, warning=FALSE, message=FALSE}
## Estimate grostrophich wind
#UV <- geostrophicwind(SLP,progressbar=FALSE)
```

```{r wind-rose}
## wind rose
#windrose(UV)
```

### 2017-04-06 EU-Circle/eu-circle_case-study-4.Rmd
Bangladesh tide measurements
Cleansed version

Bangladesh: tropical cyclones and stormsurges, heavi precipitation.

@@

```{r pressure, echo=FALSE}
# Daily precipitation from BMD 
load('bmd.rda')
```

@@

>>>>

Bangladesh: tropical cyclones and stormsurges, heavy precipitation.

@@

```{r pressure, echo=FALSE}
# Daily precipitation from BMD 
load('bmd.rda')
map(pr.bmd,new=FALSE)
```

@@

## Tidal measurements

Time series and map of the tidal data from

```{r}
load('Bangladesh-tidal-stations.rda')
plot(Z,zoom=8,new=FALSE)
```

### 2017-04-06 EU-Circle/eu-circle_case-study-5.Rmd
Cleansed version

@@

>>>>

```{r, eval=FALSE}
## Get files on the outgoing long-wave radiation (OLR).
cmip5.urls <- function(experiment='rcp45',varid='rlut',
                       url="http://climexp.knmi.nl/CMIP5/monthly/", 
                       path=NULL,off=FALSE,force=FALSE,verbose=FALSE) {
  urlfiles <- "NA"
  if(verbose) print("cmip5.urls")
  if(is.null(path)) path <- getwd()
  for (iexp in experiment) {
    if(verbose) print(iexp)
    for (ivar in varid) {
      if(verbose) print(ivar)
      ## Loop on the number of experiments
      for (irun in 0:110) { ## 
        if(verbose) print(paste(irun))
        ## Update experiment number
        if (irun < 10) run.id = paste("00",as.character(irun),sep="")
        else if (irun < 100) run.id = paste("0",as.character(irun),sep="")
        else run.id <- as.character(irun)

        ## Create output directory for the climate experiment
        path.exp <- file.path(path,experiment[grep(iexp,experiment)],
                              fsep = .Platform$file.sep)
        if (!file.exists(path.exp)) dir.create(path.exp)
        if (verbose) print(path.exp[grep(iexp,experiment)])
        ## Define the output file
        destfile <- paste(path.exp,varid[grep(ivar,varid)],sep="/") 
        destfile <- paste(destfile,"_Amon_ens_",sep="")
        destfile <- paste(destfile,iexp,sep="")
        destfile <- paste(destfile,run.id,sep="_")
        destfile <- paste(destfile,".nc",sep="")

        if (!file.exists(destfile) | force) {
          ## Update output filename with attributes:
          urlfile  <- paste(url,ivar,sep="")             # add var directory
          urlfile  <- paste(urlfile,ivar,sep="/")        # add v.name
          urlfile  <- paste(urlfile,"_Amon_ens_",sep="") # add text
          urlfile  <- paste(urlfile,iexp,sep="")         # add exp.name
          urlfile  <- paste(urlfile,run.id,sep="_")      # add exp ID number
          urlfile  <- paste(urlfile,".nc",sep="")        # add file ext
        }
        urlfiles <- c(urlfiles,urlfile)
        if (verbose) print(urlfile)
      }
    } # End for   
  }
  return(urlfiles[-1])
}

## download the CMIP5-files from the KNMI ClimateExplorer
urls <- cmip5.urls()
for (i in 1:length(urls)) if (!file.exists(substr(urls[i],43,70))) download.file(urls[i],substr(urls[i],43,70))

```

@@

A projection of future wet-day mean precipitation based on DSensemble will be carried out in two steps. First use the PCA from the original data as predictands in a set of runs with OLR as predictors. Then use the residual as estimated above as predictand for a set of DSensembles runs with the water vapour pressure as predictors. The final result is the sum of these two steps.The wet-day mean precipitation is projected through s single stemp where DSensemble uses SLP as predictor. The seasonal means are then estimated as the product between the wet-day frequency and the wet-day mean precipitation.

### 2017-04-06 EU-Circle/eu-circle_torbay-precip.Rmd
Cleansed version

>>>>

```{r, eval=FALSE}
## Get files on the outgoing long-wave radiation (OLR).
cmip5.urls <- function(experiment='rcp45',varid='rlut',
                       url="http://climexp.knmi.nl/CMIP5/monthly/", 
                       path=NULL,off=FALSE,force=FALSE,verbose=FALSE) {
  urlfiles <- "NA"
  if(verbose) print("cmip5.urls")
  if(is.null(path)) path <- getwd()
  for (iexp in experiment) {
    if(verbose) print(iexp)
    for (ivar in varid) {
      if(verbose) print(ivar)
      ## Loop on the number of experiments
      for (irun in 0:110) { ## 
        if(verbose) print(paste(irun))
        ## Update experiment number
        if (irun < 10) run.id = paste("00",as.character(irun),sep="")
        else if (irun < 100) run.id = paste("0",as.character(irun),sep="")
        else run.id <- as.character(irun)

        ## Create output directory for the climate experiment
        path.exp <- file.path(path,experiment[grep(iexp,experiment)],
                              fsep = .Platform$file.sep)
        if (!file.exists(path.exp)) dir.create(path.exp)
        if (verbose) print(path.exp[grep(iexp,experiment)])
        ## Define the output file
        destfile <- paste(path.exp,varid[grep(ivar,varid)],sep="/") 
        destfile <- paste(destfile,"_Amon_ens_",sep="")
        destfile <- paste(destfile,iexp,sep="")
        destfile <- paste(destfile,run.id,sep="_")
        destfile <- paste(destfile,".nc",sep="")

        if (!file.exists(destfile) | force) {
          ## Update output filename with attributes:
          urlfile  <- paste(url,ivar,sep="")             # add var directory
          urlfile  <- paste(urlfile,ivar,sep="/")        # add v.name
          urlfile  <- paste(urlfile,"_Amon_ens_",sep="") # add text
          urlfile  <- paste(urlfile,iexp,sep="")         # add exp.name
          urlfile  <- paste(urlfile,run.id,sep="_")      # add exp ID number
          urlfile  <- paste(urlfile,".nc",sep="")        # add file ext
        }
        urlfiles <- c(urlfiles,urlfile)
        if (verbose) print(urlfile)
      }
    } # End for   
  }
  return(urlfiles[-1])
}

## download the CMIP5-files from the KNMI ClimateExplorer
urls <- cmip5.urls()
for (i in 1:length(urls)) if (!file.exists(substr(urls[i],43,70))) download.file(urls[i],substr(urls[i],43,70))

### 2017-04-06 paper58/paper58.Rmd
Cleansed version

---
output: html_document
---
*Benestad et al - Simple and approximate estimation of future precipitation return-values* 
-----------------------------------
title: "Analysis and calculations in manuscript"
author: "Rasmus Benestad"
date: "April 4, 2016"
output: pdf_document
---

@@

```{r}
## Generate a map for the PC weights for the different modes of the
## wet-day mean mu annual cycle.
mupcamap <- function(mu,pca,ipca,xlim,ylim,r2) {
  r2 <- as.numeric(r2)
  colpc <- rev(colscal("budrd",n=100))
  cz <- round(100*abs(pca$v[,ipca])/quantile(abs(pca$v[,ipca]),0.95))
  cz[cz > 100] <- 100; cz[cz < 1] <- 1
  col <- colpc[cz]
  pch <- rep(19,length(r2))
  mo <- c(r2) > 0.4
  hi <- c(r2) > 0.6
  pch[!hi] <- 1
  pch[!mo] <- 4
  cex <- 1.2*c(r2) + 0.5
  map(mu, xlim=xlim,ylim=ylim,bg='grey90',col='grey90',cex=0.5,new=FALSE)
  points(lon(mu),lat(mu),pch=pch,col=col,cex=cex)
  par(xpd=TRUE)
  text(10,73,paste('Annual cycle in PC',ipca,
      'with variance of',round(100*pca$d[ipca]^2/sum(pca$d^2)),'%'),pos=4)

@@

print("Plot a map of projected values:")
map(mux, xlim=xlim,ylim=ylim,cex=Cex,bg='grey70',col='grey70',gridlines=FALSE,
    colbar=list(col=cols,n=12,type="p",h=0.6,v=1))
points(lon(mux),lat(mux),pch=19,col=colx,cex=Cex)
points(lon(mux),lat(mux),pch=21,col=coly,cex=Cex,lwd=2)
par(xpd=TRUE)
text(20,73,'Wet-day mean: 2100')
legend(20,32,c(expression(bar(x)),expression(q[95])),
       pch=c(21,19),bty='n',col='grey',text.col='grey',horiz=TRUE)

@@

**Figure SM1**

@@

legend(1,8.5,c(expression(bar(x)),expression(mu),expression(f[w]),
               expression(bar(n[c*w*d])),expression(bar(n[c*d*d]))),bty='n',
       col=c('steelblue','darkblue','grey','darkgreen','red'),lwd=c(3,3,1,1,1),
       ncol=2)
figlab('Figure SM1')
```

```{r}
## Figure SM2.
## test: See if the quantiles are consistent when the mean mu varies.
qtest <- aggregate(y1,year,FUN='qqexp')
qx <- c(coredata(qtest[,1:101]))
qy <- c(coredata(qtest[,102:202]))

par(bty='n')
plot(qx,qy,xlim=c(0,40),ylim=c(0,40),
     pch=19,col=rgb(0.2,0.2,0.7,0.3),cex=cex,
     main='Test: exponential distribution & changing mean',
     xlab=expression(q[p]),ylab=expression(-log(1-p)*mu))
lines(c(0,40),c(0,40),col='grey')
figlab('Figure SM2')
```

```{r}
## Show the predictor area:
X <- retrieve('air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
map(X,projection='sphere',colbar=list(pal="budrd",breaks=seq(8,28,by=0.5)))
figlab('Figure SM3',xpos=0.8,ypos=0.999)
```

@@

Figure SM7 shows the historical trends in the wet-day mean preciptation.

```{r}
## Map showing trends in mu
map(MU,FUN='trend',cex=1,colbar=list(pal="budrd",breaks=pretty(c(-0.5,0.5),n=21)))
figlab('Figure SM7',ypos=0.999)
figlab(expression(paste('Trend in ',mu,' (mm/day per decade)')),xpos=0.5,ypos=0.999)
```

@@

Figure SM9 shows the spatial distribution of the historical trends in wet-day frequency on a map.

```{r}
## Map showing trends in fw
#dev.new()
map(FW,FUN='trend',cex=1,colbar=list(pal="budrd",breaks=pretty(c(-0.03,0.03),n=21)))
figlab('Figure SM9',ypos=0.999)
figlab(expression(paste('Trend in ',f[w],' (fraction per decade)')),xpos=0.5,ypos=0.999)
```

>>>>

---
output:
  pdf_document: default
  html_document: default
---
*Benestad et al - Simple and approximate estimation of future precipitation return-values* 
-----------------------------------
title: "Analysis and calculations in manuscript"
author: "Rasmus Benestad"
date: "April 4, 2016"
output: pdf_document
---

@@

```{r}
## Generate a map for the PC weights for the different modes of the
## wet-day mean mu annual cycle.
mupcamap <- function(mu,pca,ipca,xlim,ylim,r2) {
  r2 <- as.numeric(r2)
  colpc <- rev(colscal("budrd",n=100))
  cz <- round(100*abs(pca$v[,ipca])/quantile(abs(pca$v[,ipca]),0.95))
  cz[cz > 100] <- 100; cz[cz < 1] <- 1
  col <- colpc[cz]
  pch <- rep(19,length(r2))
  mo <- c(r2) > 0.4
  hi <- c(r2) > 0.6
  pch[!hi] <- 1
  pch[!mo] <- 4
  cex <- 1.2*c(r2) + 0.5
  map(mu, xlim=xlim,ylim=ylim,bg='grey90',col='grey90',cex=0.5,gridlines=FALSE,new=FALSE)
  points(lon(mu),lat(mu),pch=pch,col=col,cex=cex)
  par(xpd=TRUE)
  text(10,73,paste('Annual cycle in PC',ipca,
      'with variance of',round(100*pca$d[ipca]^2/sum(pca$d^2)),'%'),pos=4)

@@

print("Plot a map of projected values:")
map(mux, xlim=xlim,ylim=ylim,cex=Cex,bg='grey70',col='grey70',gridlines=FALSE,
    colbar=list(col=cols,n=12,type="p",h=0.6,v=1))
points(lon(mux),lat(mux),pch=19,col=colx,cex=Cex)
points(lon(mux),lat(mux),pch=21,col=coly,cex=Cex,lwd=2)
par(xpd=TRUE)
text(20,72,'Wet-day mean: 2100')
legend(20,32,c(expression(bar(x)),expression(q[95])),
       pch=c(21,19),bty='n',col='grey',text.col='grey',horiz=TRUE)

@@

```{r}
## Figure SM1.
## test: See if the quantiles are consistent when the mean mu varies.
qtest <- aggregate(y1,year,FUN='qqexp')
qx <- c(coredata(qtest[,1:101]))
qy <- c(coredata(qtest[,102:202]))

par(bty='n')
plot(qx,qy,xlim=c(0,40),ylim=c(0,40),
     pch=19,col=rgb(0.2,0.2,0.7,0.3),cex=cex,
     main='Test: exponential distribution & changing mean',
     xlab=expression(q[p]),ylab=expression(-log(1-p)*mu))
lines(c(0,40),c(0,40),col='grey')
figlab('Figure SM1')
```

```{r}
## Show the predictor area:
X <- retrieve('air.mon.mean.nc',lon=c(-100,-30),lat=c(0,40))
map(X,projection='sphere',colbar=list(pal="budrd",breaks=seq(8,28,by=0.5)))
figlab('Figure SM2',xpos=0.8,ypos=0.999)
```

@@

legend(1,8.5,c(expression(bar(x)),expression(mu),expression(f[w]),
               expression(bar(n[c*w*d])),expression(bar(n[c*d*d]))),bty='n',
       col=c('steelblue','darkblue','grey','darkgreen','red'),lwd=c(3,3,1,1,1),
       ncol=2)
figlab('Figure SM3')
```

@@

Figure SM7 shows the historical trends in the wet-day mean preciptation. Locations with trends that are statistically significant at the 5\%-level are shown with a ring around the symbol. The significance test was based on a regression analysis and the p-value associated with the fitted slope of a linear fit.

```{r, fig.height=8}
## Map showing trends in mu
map(MU,FUN='trend',cex=1,colbar=list(pal="budrd",breaks=pretty(c(-0.5,0.5),n=21)))
pval <- apply(coredata(MU),2,'trend.pval') <= 0.05
points(lon(subset(MU,is=pval)),lat(subset(MU,is=pval)),cex=1,col=rgb(0,0,0,0.2))
figlab('Figure SM7',ypos=0.999)
figlab(expression(paste('Trend in ',mu,' (mm/day per decade)')),xpos=0.5,ypos=0.999)
```

@@

Figure SM9 shows the spatial distribution of the historical trends in wet-day frequency on a map, similar to Fig SM7.

```{r, fig.height=8}
## Map showing trends in fw
map(FW,FUN='trend',cex=1,colbar=list(pal="budrd",breaks=pretty(c(-0.03,0.03),n=21)))
pval <- apply(coredata(FW),2,'trend.pval') <= 0.05
points(lon(subset(FW,is=pval)),lat(subset(FW,is=pval)),cex=1.2,col=rgb(0,0,0,0.75))
#points(lon(FW),lat(FW),cex=1.2,col=rgb(0,0,0,0.3))
figlab('Figure SM9',ypos=0.999)
figlab(expression(paste('Trend in ',f[w],' (fraction per decade)')),xpos=0.5,ypos=0.999)
```
